\begin{abstract}
	%
	Validating wireless protocol implementations cannot always employ direct
	instrumentation of the device under test (DUT).
	%
  The DUT may not implement the required instrumentation, or the instrumentation
  may alter the DUT's behavior when enabled.
	%
	Wireless sniffers can monitor devices that direct instrumentation cannot,
	but they introduce new validation challenges.
	%
	Losses caused by wireless propagation mean that even multiple sniffers
	cannot perfectly reconstruct the actual DUT packet trace.
	%
	As a result, accurate validation using wireless sniffers requires
	distinguishing between specification deviations that represent
	implementation errors from those caused by sniffer uncertainty.

	We present a new approach that enables sniffer-based validation of wireless protocol
	implementation.
	%
  Beginning with the original protocol state machine, we automatically encode
  uncertainty introduced due to the sniffer by selectively adding
  non-deterministic transitions.
	%
	Processing the sniffer packet trace using this validation state machine
	produces a set of mutated packet traces, each of which could have been
	actually observed by the DUT.
	%
  We characterize the NP-completeness of the resulting problem and provide an
  exhaustive search algorithm for locating all mutated traces, as well as a more
  practical protocol-oblivious heuristics for locating likely mutated traces
  guided by the sniffer's packet loss characteristics.
	%
	The resulting set of mutated traces can be used to determine whether the
	protocol implementation violates the specification.
	%
	We have implemented our framework and show that it can accurately identify
	implementation errors.
	%
\end{abstract}
