\section{Prerequisites and Problem Statement}
\label{sec:background}

\subsection{Packet, Trace and Monitor State Machine}
\label{subsec:basic}

The alphabet of the monitor state machine is the finite set of all valid packets
defined by the protocol, denoted as $\mathbb{P}$.  A packet is a binary string
of finite number of bits, encoding interesting protocol attributes such as
\texttt{src}, \texttt{dest}, \texttt{type}, \texttt{flags}, and physical layer
information, such as \texttt{channel}, \texttt{modulation}, etc. The input of
the state machine then corresponds to a time-ordered sequence of packets.

\begin{definition}
  A \textit{packet trace} is a finite sequence of $(timestamp, packet)$ tuple:
  \begin{align*}
    [(t_1, p_1), (t_2, p_2),\ldots,(t_n, p_n)]
  \end{align*}%
  where $t_i \in \mathbb{Z}^+$ is the \textit{discrete} timestamp and $p_i$ is the packet
  observed at time $t_i$. The timestamps are strictly monotonically increasing:
  $t_i < t_{i+1}$ for $1 \le i < n$.
\end{definition}%

In addition to timestamp monotonicity, we also require that adjacent packets do
not overlap in time, $t_{i+1}-t_i > \text{\texttt{airtime}}(p_i)$ for $1 \le i
< n$, where \texttt{airtime()} calculates the time taken to transmit a packet.

We use \textit{timed automata}~\cite{alur1994theory} to model the expected
behaviors of the DUT.  A timed automata is a finite state machine with timing
constraints on the transitions: each transition can optionally start one or more
timers, which can latter be used to assert certain events should be seen before
or after the time out event. We refer the readers to~\cite{alur1994theory} for
more details about timed automata.


\begin{definition}
  A protocol monitor state machine $S$ is an 7-tuple\\
  $\{\Sigma, \mathbb{S}, \mathbb{X}, s_0, C, E, G\}$, where:
  \begin{itemize}
    \item $\Sigma = \mathbb{P}$ is the finite input alphabet.

    \item $\mathbb{S}$ is a non-empty, finite set of states. $s_0 \in
      \mathbb{S}$ is the initial state.

    \item $\mathbb{X}$ is the set of boolean variables. We use $v = \{x
      \leftarrow true/false\ |\ x \in \mathbb{X}\}$ to denote an assignment of
      the variables. Let $\mathbb{V}$ be the set of such values $v$. 

    \item $C$ is the set of clock variables. A \textit{clock variable} can be
      reset along any state transitions. At any instant, reading a clock
      variable returns the time elapsed since last time it was reset.

    \item $G$ is the set of guard conditions defined inductively by
      \begin{align*}
        g := true\ |\ c \le T\ |\ c \ge T\ |\ x\ |\ \neg g\ |\ g_1 \land g_2
      \end{align*}%
      where $c \in C$ is a clock variable, $T$ is a constant, and $x$ is a variable in $\mathbb{X}$. 
      A transition can choose not to use guard conditions by setting $g$ to be $true$.

    \item $E \subseteq \mathbb{S} \times \mathbb{V} \times \mathbb{S} \times
      \mathbb{V} \times \Sigma \times  G \times \mathscr{P}(C)$ gives the set of
      transitions.\\ $\langle s_i, v_i, s_j, v_j, p, g, C'\rangle \in E$
      represents that if the monitor is in state $s_i$ with variable assignments
      $v_i$, given the input tuple $(t, p)$ such that the guard $g$ is
      satisfied, the monitor can transition to a state $s_j$ with variable assignments
      $v_j$, and reset the clocks in $C'$ to 0.
  \end{itemize}
  \label{def:sm}
\end{definition}

A tuple $(t_i, p_i)$ in the packet trace means the packet $p_i$ is presented to
the state machine at time $t_i$. The monitor {\it rejects a trace} $Tr$ if there
exists a prefix of $Tr$ such that all states reachable after consuming the
prefix have no valid transitions for the next $(t, p)$ input.

As an example, the monitor state machine illustrated in
Fig.~\ref{fig:dot11_tx_ta} can be formally defined as follows:

\begin{itemize}
  \item $\Sigma = \{DATA_i, DATA'_i, Ack\ |\ 0 \le i < N\}$.
  \item Clock variables $C = \{c\}$. The only clock variable $c$ is
    used for acknowledgment time out.
  \item $\mathbb{X} = \{i\}$, as a variable with ${\mathit log}(N) + 1$ bits to count from $0$ to $N$.  
  \item Guard constraints $G = \{ c \le T_o, c > T_o, T_o < c \le T_m\}$.
    $T_o$ is the acknowledgment time out value, and $T_m >
    T_o$ is the maximum delay allowed before the retransmission packet gets
    sent. $T_o$ can be arbitrary large but not infinity in order to check the
    liveness of the DUT.
\end{itemize}


The monitor state machine defines a \textit{timed language} $L$ which consists
all valid packet traces that can be observed by the DUT.  We now give the
definition of protocol \textit{compliance} and \textit{violation}.

\begin{definition}
  Suppose $\mathbb{T}$ is the set of all possible packet traces collected from
  DUT, and $S$ is the state machine specified by the protocol. The DUT
  \textit{violates} the protocol specification if there exists an
  packet trace $Tr \in \mathbb{T}$ such that $S$ rejects $Tr$.
  Otherwise, the DUT is \textit{compliant} with the specification.
\end{definition}

The focus of this paper is to determine whether a \textit{given} $Tr$ is
evidence of a violation.
%
%We acknowledge that determining \textit{compliance} is a more challenging
%problem, as it requires enumerating every possible trace in $\mathbb{T}$, which
%is probably infinite.


\subsection{Mutation Trace}
\label{subsec:mutation}

As shown in the motivation example in Fig.~\ref{fig:sniffer_in_middle}, a
sniffer trace may either miss packets that are present in DUT trace, or contain
extra packets that are missing in DUT trace. Note that in the latter case, those
extra packets must be all sent \textit{to} the DUT. This is because it is
impossible for the sniffer to overhear packets sent from the DUT that were not
actually sent by the DUT.

We formally capture this relationship with the definition of mutation trace.

\begin{definition}
  \label{def:mutation}
  A packet trace $Tr'$ is a mutation of sniffer trace $Tr$ w.r.t a DUT if for
  all $(t, p) \in Tr\setminus Tr'$, $p.dest = DUT$.
\end{definition}

A mutation trace $Tr'$ represents a \textit{guess} of the corresponding DUT
packet trace given given sniffer trace $Tr$.  In fact, the DUT packet trace must
be one of the mutation traces of the sniffer trace $Tr$.

\begin{lemma}
  Let $Tr_{DUT}$ and $Tr$ be the DUT and sniffer packet trace captured during
  the same protocol operation session, and $\mathcal{M}(Tr)$ be the set of
  mutation traces of $Tr$ with respect to DUT, then $Tr_{DUT} \in \mathcal{M}(Tr)$.
  \label{lem:mutation}
\end{lemma}%
(Proof in Appendix~\ref{app:lemma1})


\subsection{Problem Statement}
\label{subsec:problem}

Lemma~\ref{lem:mutation} shows that $\mathcal{M}(Tr)$ is a \textit{complete} set
of guesses of the DUT packet trace. Therefore, the problem of validating DUT
implementation given a sniffer trace can be formally defined as follows:

\begin{problem}
  \label{prob:validation}
  VALIDATION\\
  \textbf{instance} A protocol monitor state machine $S$ and a sniffer trace $Tr$.\\
  \textbf{question} Does there exist a mutation trace $Tr'$ of $Tr$ that satisfies $S$?
\end{problem}
