\section{Related Works}
\label{sec:related}

\textbf{Hidden Markov Model (HMM) Approach.} When considering the whole system
under test (both DUT and endpoint), the sniffer only captures a subset of the
all the packets (events).  This is similar to the event sampling problem in
runtime
verification~\cite{bonakdarpour2011sampling,hauswirth2004low,arnold2008qvm,fei2006artemis,basin2012monitoring}.
Stoller \textit{et al}~\cite{stoller2011runtime} used HMM-based state estimation
techniques to calculate the confidence that the temporal property is satisfied
in the presence of gaps in observation.

While it seems possible to adapt the
method in~\cite{stoller2011runtime} to our problem, we note several advantages
of our augmented monitor and prioritized search procedure.  First, the
automatically augmented state machine precisely encodes the protocol
specification and the uncertainty; this is intuitive to design and naturual for
reporting the evidence for a trace being successful.  We do not require a user
to specify the number of states of underlying HMM, or accurately provide
underlying probabilities.  Second, we use timed automata to monitor the timing
constraints which are common in wireless protocols; it may be non-trivial to
encode such timing information in HMM.  Finally, we can exploit domain knowledge
to devise effective pruning heuristics to rule out unlikely sequences during the
exhaustive search. 

\textbf{Network Protocol Validation.} Lee \textit{et al}~\cite{lee1997passive}
studied the problem of passive network testing of network management. The system
input/output behavior is only partially observable. However, the uncertainty
only lies in missing events in the observation, while in the context of wireless
protocol verification, the uncertainty could also be caused by extra events not
observed by the tested system. Additionally, they do not provide any formal
guarantees even for cases when we report a definite bug.  Software model
checking techniques~\cite{musuvathi2002cmc,godefroid1997model} have also been
used to verify network protocols. Our problem is unique because of the
observation uncertainty caused by sniffers.

\begin{comment}
\textbf{Sniffer Trace Analysis.} Wireless sniffers has been widely used to
analyze MAC layer behaviors of enterprise wireless
networks~\cite{sheng:wicom2008,tan:tmc2014,yeo-wise04,yeo:witmemo2005}.
Jigsaw~\cite{Cheng:2006:JSP:1159913.1159920} is a larger scale wireless network
monitoring infrastructure. 150~radio monitors were deployed in a campus
building. Traces collected from multiple sniffers were merged and synchronized
to restructure the link and transportation layer conversations. Protocol
specific heuristics were developed to infer the missing packets. The work
in~\cite{Mahajan:2006:AMB:1159913.1159923} shared the same idea of trace merging
with Jigsaw, but uses a FSM to infer packet reception. These works assume the
correctness of the protocol implementation in order to infer missing packets,
while we systematically encode the uncertainty of sniffer traces for
verification purpose.


\textbf{ Testing Under Uncertainty}. The position paper by Rosemblum \textit{et
al}~\cite{Elbaum:2014:KUT:2635868.2666608} contains excellent motivation for the
need to combat uncertainty foundationally when testing systems. McKinley
\textit{et al}~\cite{bornholt2014uncertain,sampson2014expressing} deals with checking
assertions in programs dealing with noisy data from sensors. Instead of checking
the truth or falsity of assertions, they model the probability distribution of
the assertion conditions and perform Monte-carlo based simulations to estimate
the probabilities. Our work can be seen as leveraging non-determinism to weaken
the specification logically to precisely define the problem complexity, and use
probabilities to guide the search for likely mutations.  Other works have used
sampling to find data-race bugs~\cite{marino2009literace}, and ensure that the
sampling does not lead to spurious alarms.
\end{comment}
