\subsection{Gaming Wireless Controller Protocol}
\label{subsec:xbox}

\begin{table}[t!]
  \centering
  \begin{tabular}{lll}
    \toprule
    \textbf{Protocol Aspects} & \textbf{Traces} & \textbf{Violations (\%)}\\
    \midrule
    Sequence Number & 3049 & 1539 (50.48\%) \\
    Station Scheduling & 3046 & 2045 (67.14\%) \\
    Uplink Modulation & 3127 & 8 (0.26\%) \\
    Downlink Modulation & 3127 & 24 (0.77\%) \\
    \bottomrule
  \end{tabular}
  \caption{\textbf{Validation Results on Traces from the Gaming Wireless Protocol}}
  \label{tab:xbox}
\end{table}

We contacted a major manufacturer of gaming consoles, which uses 
%%The Xbox One controller~\cite{xbox} uses
a proprietary protocol to communicate between the gaming controllers
and the console.  The protocol is designed to achieve low-latency for
the gaming packets, and to consume low-power for longer battery life.

We obtained 75 sniffer traces from the testing team for a new version
of the protocol that is under development and testing. This team has
been testing the implementation for a few weeks.  Each trace contained
around 6 million packets that were captured during 1 hour and 40
minutes of protocol operation.

We split the traces into \num{100000} packet segments, and applied our
framework to validate the DUT implementation. We found that the latest
implementation of the protocol under development had violations of the
protocol specification. Some of the implementation bugs we found
related to the sequence number management, station scheduling, and the
changing of modulation rates adaptation. Table~\ref{tab:xbox}
summarizes the validation result. Note that if we disable the
augmented transitions, each trace will be flagged as violation because
of the missing packets, thereby reducing the usability of the tool.

The bugs related to packet sequence number and station scheduler were
detected in about half the traces, and the bug related to the rate
control algorithm was detected in only a few traces. This is because
the previous two aspects are essential in all protocol operations,
while the bugs related to rate control only manifest themselves under
certain link conditions.

After communicating with the testing team, we confirmed that the
sequence number bug was already known, as it is relatively easy to
detect even by manually examining the traces. The bug related to
station scheduling was also noticed before, yet no quantitative
results about how frequent this bug appears were obtained because of
the lack of automatized validating tools. Finally, the bug related to
rate control has been filed as a bug report.
